R for Clinical Study Reports and Submission

### 2021-09-19 tlf-ae-spec.Rmd
change patient to participants #6

Following [ICH E3
guidance](https://database.ich.org/sites/default/files/E3_Guideline.pdf),
we need to summarize which patients were included in each efficacy
analysis in Section 12.2, Adverse Events (AEs).

>>>>

Following [ICH E3
guidance](https://database.ich.org/sites/default/files/E3_Guideline.pdf),
we need to summarize which participants were included in each efficacy
analysis in Section 12.2, Adverse Events (AEs).

### 2021-09-19 tlf-disposition.Rmd
change patient to participants #6

Following [ICH E3 guidance](https://database.ich.org/sites/default/files/E3_Guideline.pdf), we need to summarize the accounting of all patients who entered the study, in Section 10.1, Disposition of Patients.
The disposition of patients reports the numbers of patients who were randomized, and who entered and completed each phase of the study, as well as the reasons for all post-randomization discontinuations, grouped by treatment and by major reason (lost to follow-up, adverse event, poor compliance, etc.).

@@

We illustrate how to prepare a report data for a simplified disposition of patients table using variables below:

@@

In the code below, we calculated patients in the analysis population.

```{r}
n_rand <- adsl %>%
  group_by(TRT01PN) %>%
  summarize(n = n()) %>%
  pivot_wider(
    names_from = TRT01PN,
    names_prefix = "n_",
    values_from = n
  ) %>%
  mutate(row = "Patients in population")
n_rand
```

In the code below, we calculate the number and percentage of patients who discontinued the study.

@@

In the code below, we calculate the number and percentage of patients who completed/discontinued the study for different reasons.

@@

In the code below, we calculate the number and percentage of patients who complete the study.
We split `n_reason` because we need to customize the row order of the table.

@@

In the code below, we calculate the number and percentage of patients who discontinued the study for different reasons.
Here we use `paste0("    ", row)` to add some leading space for individual reasons for study discontinuation.

@@

  # Table title
  rtf_title("Disposition of Patients") %>%

@@

-   Step 1: Read data into R, i.e. `adsl`.
-   Step 2: Count patients in the analysis population and name the dataset as `n_rand`.
-   Step 3: Count the number and percentage of patients who discontinued the study, and name the dataset as `n_disc`.
-   Step 4: Count the number and percentage of patients who discontinued the study in different reasons, and name the dataset as `n_reason`. Format it using `r2rtf`.
-   Step 5: Count the number and percentage of patients who completed the study, and name the dataset as `n_complete`.
-   Step 6: Bind `n_rand`, `n_disc`, `n_reason` and `n_complete` by row. Format it using `r2rtf`.

>>>>

Following [ICH E3 guidance](https://database.ich.org/sites/default/files/E3_Guideline.pdf), 
we need to summarize the accounting of all participants who entered the study, 
in Section 10.1, Disposition of Participants.
The disposition of participant reports the numbers of participants who were randomized, 
and who entered and completed each phase of the study, 
as well as the reasons for all post-randomization discontinuations, 
grouped by treatment and by major reason (lost to follow-up, adverse event, poor compliance, etc.).

@@

We illustrate how to prepare a report data for a simplified disposition of participants table using variables below:

@@

In the code below, we calculated participants in the analysis population.

```{r}
n_rand <- adsl %>%
  group_by(TRT01PN) %>%
  summarize(n = n()) %>%
  pivot_wider(
    names_from = TRT01PN,
    names_prefix = "n_",
    values_from = n
  ) %>%
  mutate(row = "Participants in population")
n_rand
```

In the code below, we calculate the number and percentage of participants who discontinued the study.

@@

In the code below, we calculate the number and percentage of participants who completed/discontinued the study for different reasons.

@@

In the code below, we calculate the number and percentage of participants who complete the study.
We split `n_reason` because we need to customize the row order of the table.

@@

In the code below, we calculate the number and percentage of participants who discontinued the study for different reasons.
Here we use `paste0("    ", row)` to add some leading space for individual reasons for study discontinuation.

@@

  # Table title
  rtf_title("Disposition of Participants") %>%

@@

-   Step 1: Read data into R, i.e. `adsl`.
-   Step 2: Count participants in the analysis population and name the dataset as `n_rand`.
-   Step 3: Count the number and percentage of participants who discontinued the study, and name the dataset as `n_disc`.
-   Step 4: Count the number and percentage of participants who discontinued the study in different reasons, and name the dataset as `n_reason`. Format it using `r2rtf`.
-   Step 5: Count the number and percentage of participants who completed the study, and name the dataset as `n_complete`.
-   Step 6: Bind `n_rand`, `n_disc`, `n_reason` and `n_complete` by row. Format it using `r2rtf`.

### 2021-09-19 tlf-efficacy.Rmd
change patient to participants #6

Following [ICH E3 guidance](https://database.ich.org/sites/default/files/E3_Guideline.pdf),
we need to summarize primary and secondary efficacy endpoints,  
in Section 11.4, Efficacy Results and Tabulations of Individual Patient.

>>>>

Following [ICH E3 guidance](https://database.ich.org/sites/default/files/E3_Guideline.pdf),
we need to summarize primary and secondary efficacy endpoints,  
in Section 11.4, Efficacy Results and Tabulations of Individual Participant.

### 2021-09-19 tlf-overview.Rmd
change patient to participants #6

A typical CSR contains full details on the methods and results of a individual clinical study.
In support of the statistical anlaysis, a large number of tables, listings and figures are
incorporated into the main text and appendices.
In CDISC pilot project, an 
[example CSR](https://bitbucket.cdisc.org/projects/CED/repos/sdtm-adam-pilot-project/browse/updated-pilot-submission-package/900172/m5/53-clin-stud-rep/535-rep-effic-safety-stud/5351-stud-rep-contr/cdiscpilot01/cdiscpilot01.pdf) 
is also provided. If you are interested in more examples of clinical study reports, 
the European Medicines Agency (EMA) publishes clinical data including 
clinical study reports submitted by pharmaceutical companies 
on [EMA clinical data website](https://clinicaldata.ema.europa.eu/web/cdp/home).

@@

- Section 10: Study patients 
- Section 11: Efficacy evaluation
- Section 12: Safety evaluation 
- Section 14: Tables, Listings and Figures referrals but not included in the text
- Section 16: Appendices

>>>>

A typical CSR contains full details on the methods and results of an individual clinical study.
In support of the statistical anlaysis, a large number of tables, listings and figures are
incorporated into the main text and appendices.
In CDISC pilot project, an 
[example CSR](https://bitbucket.cdisc.org/projects/CED/repos/sdtm-adam-pilot-project/browse/updated-pilot-submission-package/900172/m5/53-clin-stud-rep/535-rep-effic-safety-stud/5351-stud-rep-contr/cdiscpilot01/cdiscpilot01.pdf) 
is also provided. If you are interested in more examples of clinical study reports, 
the European Medicines Agency (EMA) publishes clinical data including 
clinical study reports submitted by pharmaceutical companies 
on [EMA clinical data website](https://clinicaldata.ema.europa.eu/web/cdp/home).

@@

- Section 10: Study participants 
- Section 11: Efficacy evaluation
- Section 12: Safety evaluation 
- Section 14: Tables, listings and figures referrals but not included in the text
- Section 16: Appendices

### 2021-09-19 tlf-population.Rmd
update baseline table
change patient to participants #6

---
editor_options: 
  markdown: 
    wrap: 72
---

@@

Following [ICH E3
guidance](https://database.ich.org/sites/default/files/E3_Guideline.pdf),
we need to summarize which patients were included in each efficacy
analysis in Section 11.1, Data Sets Analysed.

@@

To calculate the numbers and percentages of patients by groups, we
provide a simple (but not robust) wrapper function, `count_by()` using
the `dplyr` and `tidyr` package.

>>>>

@@

Following [ICH E3
guidance](https://database.ich.org/sites/default/files/E3_Guideline.pdf),
we need to summarize which participants were included in each efficacy
analysis in Section 11.1, Data Sets Analysed.

@@

To calculate the numbers and percentages of participants by groups, we
provide a simple (but not robust) wrapper function, `count_by()` using
the `dplyr` and `tidyr` package.
